---
title: "Adverse Events (AE)"
author: "Livia Pierotti"
date: "2026-04-01"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Adverse Events (AE)}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This tutorial demonstrates how to generate adverse event (AE) summary tables in R using the `grstat` package. AE tables are a standard tool in clinical research for evaluating the severity (grade) of events and identifying specific AE terms.

In this tutorial, it is demonstrated how to use two key functions:

-   `ae_table_grade()`, which creates table(s) summarising AE grades overall.

-   `ae_table_soc()`, which summarises adverse events by System Organ Class (SOC) and by grades. It reports the number and proportion of patients experiencing at least one event within each SOC, based on their maximum recorded AE grade.

Additional features are available via the `help()` function.

First, install packages if needed and load them.

```{r setup, message=FALSE, warning=FALSE}
library(grstat)
library(flextable)
library(dplyr)
```

# Data Overview

The function `grstat_example()`returns a set of example datasets used to demonstrate data usage.

This analysis focuses on two datasets:

-   `ae`, which contains data on adverse events.
-   `enrolres`, which includes all patients and their respective treatment arms.

The AE table provides one row for each adverse event reported for a patient, capturing key details such as the event term, its grade, the associated System Organ Class, whether it is related to the study treatment, and whether it meets the criteria for a serious adverse event. A single patient may therefore appear multiple times if they experience several events, while those with no adverse events do not appear in the table at all.

```{r}

tm = grstat_example()
attach(tm)

head(ae, 4)

head(enrolres,4)

```

# Grade Tables only: `ae_table_grade()`

Create adverse event (AE) grade tables using `ae_table_grade()`.

The function `ae_table_grade` produces AE grade table summaries based on the variant argument, which can take three different values (“max”, “sup”, “eq”) and determines the type of summary applied to AE grades. When variant = NULL (or not specified), all three summary types are returned in a single output by default.

-   The summary type, **`variant = "max"`**, generates a table showing the maximum AE grade per patient, where N is the total number of patients in the study.

-   **`variant = "sup"`**, summarises the grades greater than or equal to a specified grade per patient, where n is the number of patients having experienced at least one AE of grade ≥ X, where X ∈ {1, 2, 3, 4, 5}.

-   **`variant = "eq"`**, presents all grades for all patients. In this table, each proportion represents the number of patients who experienced a given grade at least once, where n is the number of patients having experienced at least one AE of grade equal to X, where X ∈ {1, 2, 3, 4, 5}.

## Default AE Table Grades - incuding All Variants (Max, Sup, Eq)

```{r}
ae_table_grade(data_ae=ae, data_pat=enrolres) %>% 
  as_flextable() %>% 
  fontsize(size = 8, part = "all") %>% 
  padding(padding.top=0, padding.bottom=0)
```

## Custom table of maximum AE grade per patient per arm

As an extra example of AE grade table, a table of maximum grade per patient and per arm (i.e. variant = "max") is also presented.

```{r}
ae_table_grade(data_ae=ae, data_pat=enrolres, arm="arm", variant="max") %>% 
  as_flextable() %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated per given arm.") %>% 
  fontsize(size = 8, part = "all") %>% 
  padding(padding.top=0, padding.bottom=0)
```

# SOC's tables: `ae_table_soc()`

The function ae_table_soc() creates a summary table of AE grades for each patient according to AE term and SOC Common Terminology Criteria for Adverse Events. (CTCAE). The resulting dataframe can be piped to as_flextable() to get a nicely formatted flextable.

## Table of AE grades by SOC

This table summarises the distribution of adverse event grades by System Organ Class (SOC). You can add total = FALSE to remove the “Total” column.

```{r}
ae_table_soc(df_ae=ae, df_enrol=enrolres, arm=NULL, term=NULL, 
             sort_by_count=FALSE) %>% 
  head(10) %>% 
  as_flextable() %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated.") 
```

## Table of SOC and AE terms, all grades combined, stratified by arm

Presents SOC and Preferred Term frequencies with all AE grades combined, stratified by treatment arm. Provides an overall summary of event occurrence per arm, regardless of grade.

```{r}
ae_table_soc(df_ae=ae, df_enrol=enrolres, arm="arm", term="aeterm", 
             sort_by_count=FALSE) %>%
  head(10) %>% 
  as_flextable() %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated.")
```

**Using sort_by_count=TRUE**

```{r}
ae_table_soc(df_ae=ae, df_enrol=enrolres, arm="arm", term="aeterm", 
             sort_by_count=TRUE) %>%
  as_flextable() %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated.")
```
