---
title: "Adverse Events (AE)"
author: "Livia Pierotti"
date: "2026-04-01"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Adverse Events (AE)}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This tutorial demonstrates how to generate adverse event (AE) summary tables in R using the `grstat` package. AE tables are a standard tool in clinical research for evaluating the severity (grade) of events and identifying specific AE terms.

In this tutorial, it is demonstrated how to use two key functions:

-   `ae_table_grade()`, which creates table(s) summarising AE grades overall.

-   `ae_table_soc()`, which summarises adverse events by System Organ Class (SOC) and by grades. It reports the number and proportion of patients experiencing at least one event within each SOC, based on their maximum recorded AE grade.

Additional features are available via the `help()` function.

First, install packages if needed and load them.

```{r setup, message=FALSE, warning=FALSE}
library(grstat)
library(flextable)
library(dplyr)
```

# Data Overview

The function `grstat_example()`returns a set of example datasets used to demonstrate data usage.

This analysis focuses on two datasets:

-   `ae`, which contains data on adverse events.
-   `enrolres`, which includes all patients and their respective treatment arms.

The AE table provides one row for each adverse event reported for a patient, capturing key details such as the event term, its grade, the associated System Organ Class, whether it is related to the study treatment, and whether it meets the criteria for a serious adverse event. A single patient may therefore appear multiple times if they experience several events, while those with no adverse events do not appear in the table at all.

```{r}

tm = grstat_example()
attach(tm, warn.conflicts=FALSE)

head(ae, 4)

head(enrolres,4)

```

# Grade Tables only: `ae_table_grade()`

The function `ae_table_grade()` summarises toxicity data at the patient level.

The `measure` argument controls how AE grades are summarised:

- **`measure = "max"`** summarises the worst reported AE grade observed for each patient. Rows are mutually exclusive.
- **`measure = "sup"`** summarises cumulative grade thresholds. For example, `Grade ≥ 3` counts patients with at least one AE of grade 3, 4, or 5.
- **`measure = "eq"`** summarises exact-grade occurrence. For example, `Grade 3` counts patients with at least one grade 3 AE.

Percentages are calculated using the total number of patients as denominator.

By default, all three summaries are returned in a single table. To return only one summary, specify the corresponding `measure` value.

Use `as_flextable()` to format the output as a `flextable` and add footnotes or other formatting as needed. 

Use cases are described below. See `help(ae_table_grade)` for the full list of arguments and some more examples.

## Default: AE Table Grades - incuding All Variants (Max, Sup, Eq)

```{r}
ae_table_grade(data_ae=ae, data_pat=enrolres) %>% 
  as_flextable() %>% 
  fontsize(size = 8, part = "all") %>% 
  padding(padding.top=0, padding.bottom=0)
```

## Alternative: Maximum AE grade per patient per arm

As an extra example of AE grade table, a table of maximum grade per patient and per arm (i.e. variant = "max") is also presented.

```{r}
ae_table_grade(data_ae=ae, data_pat=enrolres, arm="arm", variant="max") %>% 
  as_flextable() %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated per given arm.") %>% 
  fontsize(size = 8, part = "all") %>% 
  padding(padding.top=0, padding.bottom=0)
```

# SOC's tables: `ae_table_soc()`

The function `ae_table_soc()` summarises AE grades by event group at the patient level.

The main grouping variable is defined by `group1`, which is typically the System Organ Class (SOC). An optional second grouping variable, `group2`, can be used for more detailed summaries, for example by preferred term within SOC.

It uses the same `measure` argument as `ae_table_grade()`, but only one is used. By default, `measure = "max"`, as this is the most common summary for safety tables. For each group, patients are then classified according to the maximum reported AE grade observed within that group.

The resulting table contains one row per `group1` level and one column per grade. The `tot` column reports the number of patients with at least one AE in the corresponding group. The `na` column reports patients with at least one AE in the group but no available grade for the maximum-grade assessment.

The resulting data frame can be piped to `as_flextable()` to produce a formatted table. In the header, `N` is the total number of patients. Percentages are calculated using this total as denominator.

## Default: AE maximum grades by SOC

```{r}
ae_table_soc(data_ae=ae, data_pat=enrolres, group1="aesoc") %>% 
  as_flextable() %>% 
  delete_rows(i=11:27) %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated.") 
```

## Alternative: AE maximum grades by SOC and term, stratified by arm

```{r}
ae_table_soc(data_ae=ae, data_pat=enrolres, arm="arm", group1="aesoc", group2="aeterm") %>%
  as_flextable() %>% 
  delete_rows(i=11:165) %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients whose maximum AE grade was as indicated.")
```

## Alternative 2: AE minimum grades by SOC and term, stratified by arm

The function handles many more arguments. For example, you can use `sort_by_count=FALSE` to sort alphabetically by `group1` instead of by count, and `total=FALSE` and `showNA=FALSE` to remove the "Total" and "NA" columns.

```{r}
ae_table_soc(data_ae=ae, data_pat=enrolres, measure="sup", arm="arm", group1="aeterm", 
             sort_by_count=FALSE, total=FALSE, showNA=FALSE) %>%
  as_flextable() %>% 
  delete_rows(i=11:104) %>% 
  add_footer_lines("In the header, N represents the number of patients.") %>% 
  add_footer_lines("Percentages reflect the proportion of patients who had at least one AE higher or equal to the indicated grade.")
```
