---
title: "Calculation of RECIST Best Overall Response in grstat"
author: "Alexis COCHARD"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Calculation of RECIST Best Overall Response in grstat}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette illustrates the calculation of RESIST endpoints in R using `grstat` package.

## RECIST definition

RECIST (Response Evaluation Criteria in Solid Tumors) is a standardized way to quantify how solid tumors respond to therapy by comparing changes in tumor burden on imaging. The four possible responses are:

-   Complete Response (CR) : disappearance of all target lesions,
-   Partial Response (PR) : ≥ 30% decrease in the sum of longest diameters from baseline,
-   Stable Disease (SD) : neither sufficient shrinkage for PR nor sufficient increase for progression,
-   Progressive Disease (PD) : ≥ 20% increase in the sum from the smallest measured value (the “nadir”) or appearance of new lesions.

We follow the RECIST 1.1 guideline, DOI: 10.1016/j.ejca.2008.10.026 available [here](https://project.eortc.org/recist/wp-content/uploads/sites/4/2015/03/RECISTGuidelines.pdf)

## Endpoint definition

The Best Overall Response (BOR) is defined as the best response observed for a patient during follow-up. BOR captures the maximum tumor shrinkage achieved under treatment, reflecting the patient’s best observed benefit, although it does not account for its duration.

Several summary endpoints are then defined as proportions derived from BOR:

-   Objective Response Rate (ORR): Proportion of patients with BOR equal to complete response (CR) or partial response (PR).
-   Clinical Benefit Rate (CBR): Proportion of patients with BOR equal to CR, PR, or stable disease (SD) lasting at least a predefined duration (e.g., 6 months).
-   Disease Control Rate (DCR): Proportion of patients with BOR equal to CR, PR, or SD.

Note that ORR is sometimes defined at a specific timepoint rather than based on BOR. Such definitions depend on follow-up duration and may introduce ambiguity due to censoring and incomplete assessments

## Confirmed and Unconfirmed response

Most of the time, BOR is calculated based on the unconfirmed response, but some time a confirmed response is required. We defined the Unconfirmed and confirmed response as follow.

-   An unconfirmed best response is defined as the best categorical response observed from treatment start until progression or end of follow-up
-   A confirmed best response is defined as the best categorical consecutive response (with a minimum delay of 28 days) observed from treatment start until progression or end of follow-up. For example :
    -   A CR-CR-SD-PD consecutive response is a confirmed CR best response;

    -   A PR-SD-SD-PR is a confirmed SD best response;

    -   A single NE (and a single SD if you follow the [PharmaSUG recommendation](https://pharmasug.org/proceedings/2023/QT/PharmaSUG-2023-QT-047.pdf)) between two CR or two PR doesn't prevent confirmation.

See RECIST 1.1 guidelines for more information

# Data Preparation

To calculate BOR, we need a RECIST dataset in long format (one row per patient assessment) containing at least the following columns :

-   `subjid`, Patient ID
-   `rcdt`, Evaluation date
-   `rcresp`, Global response
-   `rctlsum`, Sum of target lesions

In this vignette, we use `grstat_example()` to simulate such data.

Note that warnings about your RECIST data may appear, so be sure to take them into account!

```{r, include=FALSE}
options(grstat_lifecycle_verbosity="quiet")
```

```{r message=FALSE}
library(grstat)
library(tidyverse)
library(flextable)

db = grstat_example(N=200)
recist = db$recist %>% select(subjid, rcdt, rctlsum, rcresp)
recist %>% filter(subjid < 4)
```

# Calculation of BOR

The calculation of the Best Overall Response is performed using `calc_best_response()`, which takes N parameters :

-   `data_recist`, the recist data.frame
-   `confirmed`, for a confirmed or unconfirmed response
-   `cols`, a vector with column names (`subjid`, `rc_resp`, `rc_date`, and `rc_sum`, the column names for, respectively, Patient ID, Global response, Evaluation date, and Sum of target lesions)

## Unconfirmed response

```{r}
best_resp = calc_best_response(recist)
head(best_resp, n= 7L)
```

As you can see, there are warnings. You can go to `check_recist()` for more information.

If your column names aren't recognized, you can specify them using the parameter cols

```{r eval=FALSE}
best_resp = calc_best_response(recist, cols = c(rc_resp="Glb_resp", subjid="Pat_Num"),)

```

## Confirmed response

```{r}
confirmed_best_resp = calc_best_response(recist, confirmed = TRUE)
head(confirmed_best_resp, n = 7L)

#To illustrate the difference between a confirmed and an unconfirmed response, let's focus on patient 7.
recist %>% filter(subjid ==7)

```

With the confirmed method, patient 7 best response becomes Stable disease instead of Partial response. The reason is that patient 7 has a Progressive disease after his Partial response. Following the definition, the Partial response can't be confirmed. That's why the confirmed best response is now a Stable disease instead of a Partial response for the unconfirmed best response

# Calculate the different RECIST rates

To summarize patient-level responses and compute derived rates, we use the `aggregate_recist_rates()` function that consist of 2 paramater :

-   `data`, the data.frame containing the best response obtain after using the function `calc_best_response()` (confirmed or unconfirmed)
-   `derived_endpoints=c("ORR", "CBR", "DCR")`, to show all the endpoint we want

```{r}
tbl = aggregate_recist_rates(best_resp, derived_endpoints=c("ORR", "CBR", "DCR"))
tbl
```

As `aggregate_recist_rates()` returns a classed data frame, we can use `as_flextable()` (instead of simply `flextable()`) to turn it into a formatted flextable, on which you can apply any modifier. Here, for example, we apply bold on row 4 and add a footnote on the 3rd row of the CI column.

```{r}
#Exampel with flextable modification:
tbl %>% 
  as_flextable() %>% 
  bold(i=4)%>%
  footnote(i = 3, j = 4,
                value = as_paragraph("This is note number 8"),
                ref_symbols ="8", part = "body")
```
