---
title: "Guideline Suggestions for the calculation of RESIST endpoints"
author: "Alexis COCHARD"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Guideline Suggestions for the calculation of RESIST endpoints}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette illustrates the calculation of RESIST endpoints in R using `grstat` package.

## RECIST definition

RECIST (Response Evaluation Criteria in Solid Tumors) is a standardized way to quantify how solid tumors respond to therapy by comparing changes in tumor burden on imaging. The four possible responses are:

-   Complete Response (CR) (disappearance of all target lesions),
-   Partial Response (PR) (≥ 30% decrease in the sum of longest diameters from baseline),
-   Stable Disease (SD) (neither sufficient shrinkage for PR nor sufficient increase for progression),
-   Progressive Disease (PD) (≥ 20% increase in the sum from the smallest measured value (the “nadir”) or appearance of new lesions).

We follow the RECIST 1.1 guideline, DOI: 10.1016/j.ejca.2008.10.026

## Endpoint definition

Defenition of different endpoint linked to the RECIST evaluation and present in the grstat package :

-   Best Overall Response (BOR): Measured the proportion of patients achieving a complete or partial response to treatment within a specified time frame. In a clinical trial, measuring the BOR is one way to see how well a new treatment works.
-   Clinical Benefit Rate (CBR): Proportion of complete response, partial response, and stable disease lasting at least six months to assess overall treatment benefit.
-   Disease Control Rate (DCR) : Proportion of complete response, partial response, and stable disease.

## Confirmed and Unconfirmed response

Most of the time, BOR is calculated based on the unconfirmed response, but some time a confirmed response is required. We defnied the Unconfirmed and confirmed response as follow.

-   An unconfirmed best response is defined as the best categorical response observed from treatment start until progression or end of follow-up
-   A confirmed best response is defined as the best categorical consecutive response (with a minimum delay of 28 days) observed from treatment start until progression or end of follow-up. For example a CR-CR-SD-PD consecutive response is a confirmed CR best response while a PR-SD-SD-PR is a confirmed SD best response. A single SD or NE between two CR or two PR doesn't prevent confirmation. See RECIST 1.1 guidelines for more information

# Data Preparation

To calculate the BOR, we'll use the "recist" dataset. This dataset must be in long format (one line represents a single measure of a patient's recist criterion) and must contain at least the columns :

-   `subjid`, Patient ID
-   `rcdt`, Evaluation date
-   `rcresp`, Global response
-   `rctlsum`, Sum of target lesions

We use `grstat_example()` to simulate data.

Warnings may appear, take them into account when using the various functions

```{r, include=FALSE}
options(grstat_lifecycle_verbosity="quiet")
```

```{r}
library(grstat)
library(tidyverse)
library(flextable)

db = grstat_example(N=200)
recist = db$recist %>% select(subjid, rcdt, rctlsum, rcresp)
recist %>% filter(subjid < 4)
```

# Calculation of BOR

`calc_best_response()` consist of 6 paramater :

-   `data_recist`, the recist data.frame
-   `subjid`, the column for Patient ID
-   `rc_resp`, the column for Global response
-   `rc_date`, the column for Evaluation date
-   `rc_sum`, the column for the Sum of target lesions
-   `confirmed`, for a confirmed or unconfirmed response \## Unconfirmed response

```{r eval=FALSE}
#To change the colonnes names
best_resp = calc_best_response(recist, subjid="test", rc_resp="test 2")

```

```{r}
best_resp = calc_best_response(recist)
head(best_resp, n= 7L)

```

As you can see, there are warnings. You can go to `check_recist()` for more information.

## Confirmed response

```{r}
confirmed_best_resp = calc_best_response(recist, subjid="subjid", rc_resp="rcresp", 
                                         rc_date="rcdt", rc_sum = "rctlsum", confirmed = TRUE)
head(confirmed_best_resp, n = 7L)

recist %>% filter(subjid ==7)

```

With the confirmed method, patient 7 best response becomes Stable disease instead of Partial response. The reason is that patient 7 has a Progressive disease after his Partial response. Following the definition, the Partial response can't be confirmed. That's why the confirmed best response is now a Stable disease instead of a Partial response for the unconfirmed best response

# Cration of a dataframe for BOR and CBR

`aggregate_recist_rates()` consist of 2 paramater :

-   `data`, the data.frame containing the best response obtain after using the function `calc_best_response()` (confirmed or unconfirmed)
-   `derived_endpoints=c("BOR", "CBR", "DCR")`, to show all the endpoint we want

As `aggregate_recist_rates()` returns a classed data frame, we can use `as_flextable()` (instead of simply `flextable()`) to turn it into a formatted flextable, on which you can apply any modifier. Here, for example, we apply bold on row 4 and add a footnote on the 3rd row of the CI column.

```{r}
tbl = aggregate_recist_rates(best_resp, derived_endpoints=c("BOR", "CBR", "DCR"))
tbl

#Exampel with flextable modification:
tbl2 = aggregate_recist_rates(confirmed_best_resp, derived_endpoints=c("BOR", "CBR", "DCR")) %>% 
  as_flextable() %>% 
  bold(i=4)%>%
  footnote(i = 3, j = 4,
                value = as_paragraph("This is note number 8"),
                ref_symbols ="8", part = "body")
tbl2
```
