---
title: "wip_ae_grade"
output: 
  pdf_document:
    toc: true
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{wip_ae_grade}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center", 
  fig.show="hold", 
  fig.width=7,
  fig.height=5
)
```

```{r setup}
# devtools::load_all()
library(grstat)
```

## Same effect

```{r, fig.show='hold'}
tm = grstat_example(N=500, n_max=50, p_sae=0.1, beta_trt=0, beta0=-1)
ae_plot_grade(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_ridges(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
```

## Higher AE grade (same amount)

The experimental treatment causes the same amount of AE, but they are more likely to be Serious AE (20% vs 10%) and of higher grade (beta=0.3).

```{r}
tm = grstat_example(N=502, n_max=50, n_max_trt=50, p_sae=0.1, p_sae_trt=0.2, beta_trt=0.3)
ae_plot_grade(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_ridges(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
```

In this case especially, it is interesting to weight the grades, so that higher grades count more on the Y axis.

```{r}
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM", weights=c(1,1,3,6,10))
```


## Higher AE amount (same grade)

The experimental treatment causes a higher amoung of AE, but they are as likely to be Serious AE (10%) and on average of same grade (beta=0).

```{r}
tm = grstat_example(N=504, n_max=50, n_max_trt=100, beta_trt=0)
ae_plot_grade(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_ridges(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
```

In this case, weighting the grades is not very useful.

```{r}
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM", weights=c(1,1,3,6,10))
```


## Exp treatment is globally more toxic

The experimental treatment causes more AE (max 70 vs 50), which are more likely to be Serious AE (20% vs 10%). It also gives AE of higher grade (beta=0.3).

```{r}
tm = grstat_example(N=506, p_sae=0.1, p_sae_trt=0.2, n_max=50, n_max_trt=70, beta_trt=0.3)
ae_plot_grade(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_sum(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
ae_plot_grade_ridges(df_ae=tm$ae, df_enrol=tm$enrolres, arm="ARM")
```
