---
title: "RECIST Check"
output: 
  html_document:
    toc: true
params:
  recist_check: NULL
---

```{r init, include=FALSE}
library(dplyr)
library(flextable)

knitr::opts_chunk$set(echo = FALSE)


flextable2 = function(x, autofit=TRUE, fontsize_body=8, fontsize_head=9, padding_v=0, ...){
  x = flextable(x, ...)
  if(autofit) x = set_table_properties(x, layout="autofit")
  x %>% 
    fontsize(size=fontsize_body) %>% 
    fontsize(size=fontsize_head, part="head") %>% 
    padding(padding.top=padding_v, padding.bottom=padding_v) %>% 
    set_header_labels(message = "Message",
                      n_subjid = "Number of patients")
}

show_summaries = function(x){
  sup0 = which(x$n_subjid>0)
  for(i in sup0){
    d = x %>% slice(i)
    cat(glue("<details><summary><h5 style='display:inline-block'> {d$message} (N = {d$n_subjid} patients)</h5></summary>\n\n"))
    print(htmltools::tagList(DT::datatable(d$data[[1]])))
    cat("</details>\n\n")
  }
}

recist_check = params$recist_check
rc_errors = as_tibble(recist_check) %>% 
  filter(level=="ERROR")
rc_warnings = as_tibble(recist_check) %>% 
  filter(level=="WARNING")
rc_checks = as_tibble(recist_check) %>% 
  filter(level=="CHECK")

```

<button id="toggleButton" onclick="toggleDetails()">Open all </button>


# Errors

These checks signal errors that should be corrected in the CRF.

```{r}
rc_errors %>% 
  select(message, n_subjid) %>% 
  mutate(n_subjid = ifelse(n_subjid==0, "\U2713", n_subjid)) %>% 
  flextable2()
```

```{r, results='asis'}
show_summaries(rc_errors)
```


# Warnings

These checks signal issues that can be problematic, but that can be worked around.

```{r}
rc_warnings %>% 
  select(message, n_subjid) %>% 
  mutate(n_subjid = ifelse(n_subjid==0, "\U2713", n_subjid)) %>% 
  flextable2()
```

```{r, results='asis'}
show_summaries(rc_warnings)
```

# Checks

These checks signal cases that are possible but should probably be confirmed.

```{r}
rc_checks %>% 
  select(message, n_subjid) %>% 
  mutate(n_subjid = ifelse(n_subjid==0, "\U2713", n_subjid)) %>% 
  flextable2()
```

```{r, results='asis'}
show_summaries(rc_checks)
```


```{r, results='asis'}
#I don't know why, but rendering fails if this is removed...
DT::datatable(tibble())
```


# Appendix

## Mapping

The following mapping table was used to map the checks and the RECIST dataset's columns.

```{r}
recist_check %>% 
  attr("mapping") %>% 
  tibble::enframe() %>% 
  flextable2() %>% 
  set_header_labels(name="Check column", value="Dataset column")
```


<script>
function toggleDetails() {
    const detailsElements = document.querySelectorAll("details");
    const button = document.getElementById("toggleButton");
    
    const allOpen = Array.from(detailsElements).every(details => details.open);
    
    detailsElements.forEach(details => {
        details.open = !allOpen; // Si tout est ouvert, on ferme, sinon on ouvre
    });
    
    button.textContent = allOpen ? "Open all" : "Close all";
}
</script>
